---
# for rapid prototyping:
output: bookdown::html_document2
# for final pdf:
# output:
#   bookdown::pdf_document2:
#     template: GISRUKPaperTemplate.tex
bibliography: references.bib
title: "Disaggregating origin-destination data: methods, implementations, and optimal parameters for generating accurate route networks for sustainable transport planning"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE, eval=FALSE}
rbbt::bbt_update_bib("README.Rmd", path_bib = "references.bib")
# count words
wordcountaddin::text_stats_chr(readLines("README.Rmd"))
# get the data:
system("wget https://github.com/dabreegster/odjitter/archive/refs/heads/main.zip")
unzip("main.zip")
file.rename("odjitter-main/data", "data")
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  cache = TRUE
)
```

# Introduction

Origin-destination (OD) datasets are used to represents movement through geographic space, from an origin (O) to a destination (D).
Also referred to as 'flow data'^[https://www.ons.gov.uk/census/2011census/2011censusdata/originanddestinationdata] OD datasets usually contain not only information about where they start or end, but also about the amount of movement between zones (which are often represented by a zone centroid) or other geographic entities.
Because of their ability to encode a large amount of information about millions of trips in a relatively small amount of storage space, with the maximum number of rows in an aggregate OD dataset equal to square of the number of zones squared, including intra-zonal OD pairs.
Thus, the entire transport system of London can be represented, albeit simplistically, as an OD dataset representing movement between the city's 33 boroughs with only 33^2 (1089) rows and a number of columns depending on the number of trip types.

Because of these easy-to-use characteristics, OD datasets have long been used to describe aggregate urban mobility patterns [@carey_method_1981].
Typically, and historically to reduce computational requirements, the *geographic* representation of OD datasets has been based on zone centroids, with all trips shown as departing from and arriving to a single centroid per zone.
Connecting zones with straight lines in this way results in geographic 'desire lines' which can be converted into geographic objects or simply used for visualisation purposes.
This approach to the geographic visualisation of OD data makes sense: it can provide a quick summary of the movement patterns between zones, particularly when attributes such as total number of trips and the percentage of trips made by a particular mode are represented by variable aesthetics such as line width and colour.

<!-- Todo: add figure showing this? -->

# Data and methods

# Findings


# Discussion


Acknowledgements
================

Acknowledgement should be made of any funding bodies who have supported
the work reported in the paper, of those who have given permission for
their work to be reproduced or of individuals whose particular
assistance is due recognition. Acknowledge data providers here where
appropriate.

Biography
=========

All contributing authors should include a biography of no more than 50
words each outlining their career stage and research interests.

# References
